---
title: "BSHES 740 Assignment 1"
author: "JaMor Hairston"
date: "2024-10-4"
output: html_document
---

```{r}
library(dplyr)

# Load datasets
A1_set1 <- read.csv("TADA_A1_set1.csv")
A1_set2 <- read.csv("TADA_A1_set2.csv")

#View the first 5 tweets
head(A1_set1, 5)
head(A1_set2, 5)
```


1. What are the date ranges for the two sets? What information is provided in the CSV files? What are the languages in which tweets have been posted? (2 points)

```{r}
# Display information about the datasets and the length. (Question 2)
summary(A1_set1)
summary(A1_set2)
```

```{r}
# Convert the 'date' column to Date type
A1_set1$date <- as.Date(A1_set1$date, format="%m/%d/%y")

# Find the date range
date_range <- range(A1_set1$date, na.rm = TRUE)
print(paste("A1_set1 Date range:", date_range[1], "to", date_range[2]))

# Get the unique languages
languages <- unique(A1_set1$lang)
print("Languages in which A1_set1 tweets have been posted:")
print(languages)
```

```{r}
# Convert the 'date' column to Date type
A1_set2$date <- as.Date(A1_set2$date, format="%Y/%m/%d")

# str(A1_set2$date)

# Find the date range
date_range <- range(A1_set2$date, na.rm = TRUE)
print(paste("A1_set2 Date range:", date_range[1], "to", date_range[2]))

# Get the unique languages
languages <- unique(A1_set2$lang)
print("Languages in which A1_set2 tweets have been posted:")
print(languages)
```

2. How many tweets are there for methadone, Suboxone, and fentanyl in total?

```{r}
# Convert the text column to lowercase for easier pattern matching
A1_set1$text <- tolower(A1_set1$text)
A1_set2$text <- tolower(A1_set2$text)

# Define substances keywords
patterns <- c("methadone", "suboxone", "bupes", "fentanyl", "fent")

# Filter rows containing any of the substances in the text
substance_tweets1 <- A1_set1 %>%
  filter(grepl(paste(patterns, collapse = "|"), text))

substance_tweets2 <- A1_set2 %>%
  filter(grepl(paste(patterns, collapse = "|"), text))

# Count the total number of tweets containing these substances
total_substance_tweets1 <- nrow(substance_tweets1)
total_substance_tweets2 <- nrow(substance_tweets2)
print(paste("Total number of A1_set1 tweets for key substances:", total_substance_tweets1))
print(paste("Total number of A1_set2 tweets for key substances:", total_substance_tweets2))
```

4. Are there fentanyl analogs that are also being discussed (eg., carfentanil)?

```{r}
fent_analogs <- c("carfentanil", "acetylfentanyl", "furanylfentanyl", "cyclopropylfentanyl", "acrylfentanyl")

# Search for posts that mention fentanyl analogs
analog_posts1 <- A1_set1 %>%
  filter(grepl(paste(fent_analogs, collapse = "|"), text))

analog_posts2 <- A1_set2 %>%
  filter(grepl(paste(fent_analogs, collapse = "|"), text))

# View the filtered posts mentioning fentanyl analogs
print(paste("Number of fentanyl analogs A1_set1 posts :", nrow(analog_posts1)))
print(paste("Number of fentanyl analogs A1_set2 posts :", nrow(analog_posts2)))
```

5. What are some of the topics that are most closely associated with each of the three substances?

```{r}
library(tidyr)
library(tidytext)
library(ggplot2)

# Define substances and words to ignore
substances <- c("methadone", "suboxone", "fentanyl")
ignore_words <- c("https", "t.co", "methadone_cat", "amp", "danlairdmd", "tootwisted40", "2")

# Function to clean and tokenize text
clean_and_tokenize <- function(dataset, substance) {
  dataset %>%
    filter(grepl(substance, text)) %>%              # Filter for posts mentioning the substance
    unnest_tokens(word, text) %>%                   # Tokenize the text
    filter(!word %in% ignore_words) %>%             # Remove ignored words
    filter(!word %in% stop_words$word) %>%          # Remove common stop words
    filter(word != substance) %>%                   # Exclude the substance name itself
    count(word, sort = TRUE) %>%                    # Count word frequencies
    top_n(10, n)                                    # Return the top 10 most frequent words
}

# Load the default stop words from tidytext
data(stop_words)

# Get the top words for each substance
top_methadone_s1 <- clean_and_tokenize(A1_set1, "methadone")
top_suboxone_s1 <- clean_and_tokenize(A1_set1, "suboxone")
top_fentanyl_s1 <- clean_and_tokenize(A1_set1, "fentanyl")


print(top_methadone_s1)
print(top_suboxone_s1)
print(top_fentanyl_s1)

# Visualize the top topics for each substance
ggplot(top_methadone_s1, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top A1_set1 Topics Associated with Methadone", x = "Words", y = "Frequency")

ggplot(top_suboxone_s1, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top A1_set1 Topics Associated with Suboxone", x = "Words", y = "Frequency")

ggplot(top_fentanyl_s1, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top A1_set1 Topics Associated with Fentanyl", x = "Words", y = "Frequency")

```

```{r}
# Get the top words for each substance
top_methadone_s2 <- clean_and_tokenize(A1_set2, "methadone")
top_suboxone_s2 <- clean_and_tokenize(A1_set2, "suboxone")
top_fentanyl_s2 <- clean_and_tokenize(A1_set2, "fentanyl")

print(top_methadone_s2)
print(top_suboxone_s2)
print(top_fentanyl_s2)

# Visualize the top topics for each substance
ggplot(top_methadone_s2, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top A1_set2 Topics Associated with Methadone", x = "Words", y = "Frequency")

ggplot(top_suboxone_s2, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top A1_set2 Topics Associated with Suboxone", x = "Words", y = "Frequency")

ggplot(top_fentanyl_s2, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top A1_set2 Topics Associated with Fentanyl", x = "Words", y = "Frequency")
```

6. Generate word clouds for each set.

```{r}
library(wordcloud)

# Generate word clouds for each substance
set.seed(1234)

# Methadone
wordcloud(words = top_methadone_s1$word, freq = top_methadone_s1$n, min.freq = 1,
          random.order = FALSE, rot.per = 0.35, colors = brewer.pal(8, "Dark2"), scale = c(3, 0.5))
title("Word Cloud for A1_set1 Methadone")

# Suboxone
wordcloud(words = top_suboxone_s1$word, freq = top_suboxone_s1$n, min.freq = 1,
          random.order = FALSE, rot.per = 0.35, colors = brewer.pal(8, "Set3"), scale = c(3, 0.5))
title("Word Cloud for A1_set1 Suboxone")

# Fentanyl
wordcloud(words = top_fentanyl_s1$word, freq = top_fentanyl_s1$n, min.freq = 1,
          random.order = FALSE, rot.per = 0.35, colors = brewer.pal(8, "Blues"), scale = c(3, 0.5))
title("Word Cloud for A1_set1 Fentanyl")

```

```{r}
library(wordcloud)

# Generate word clouds for each substance
set.seed(1234)

# Methadone
wordcloud(words = top_methadone_s2$word, freq = top_methadone_s2$n, min.freq = 1,
          random.order = FALSE, rot.per = 0.35, colors = brewer.pal(8, "Dark2"), scale = c(3, 0.5))
title("Word Cloud for A1_set2 Methadone")

# Suboxone
wordcloud(words = top_suboxone_s2$word, freq = top_suboxone_s2$n, min.freq = 1,
          random.order = FALSE, rot.per = 0.35, colors = brewer.pal(8, "Set3"), scale = c(3, 0.5))
title("Word Cloud for A1_set2 Suboxone")

# Fentanyl
wordcloud(words = top_fentanyl_s2$word, freq = top_fentanyl_s2$n, min.freq = 1,
          random.order = FALSE, rot.per = 0.35, colors = brewer.pal(8, "Blues"), scale = c(3, 0.5))
title("Word Cloud for A1_set2 Fentanyl")
```

7. Generate appropriate time-series figures to compare how the frequencies of mentions of these substances differ.

```{r}
# Filter for tweets mentioning each substance and count by date
S1_mentions_by_date <- A1_set1 %>%
  mutate(methadone = grepl("methadone", text),
         suboxone = grepl("suboxone", text),
         fentanyl = grepl("fentanyl", text)) %>%
  group_by(date) %>%
  summarise(methadone_mentions = sum(methadone),
            suboxone_mentions = sum(suboxone),
            fentanyl_mentions = sum(fentanyl))

# Reshape data for plotting
mentions_long <- S1_mentions_by_date %>%
  pivot_longer(cols = c(methadone_mentions, suboxone_mentions, fentanyl_mentions),
               names_to = "substance", values_to = "mentions")

# Plot the time-series comparison
ggplot(mentions_long, aes(x = date, y = mentions, color = substance)) +
  geom_line(size = 1) +
  labs(title = "A1_set1 Mentions of Methadone, Suboxone, and Fentanyl Over Time",
       x = "Date", y = "Number of Mentions",
       color = "Substance") +
  theme_minimal()
```

```{r}
# Filter for tweets mentioning each substance and count by date
S2_mentions_by_date <- A1_set2 %>%
  mutate(methadone = grepl("methadone", text),
         suboxone = grepl("suboxone", text),
         fentanyl = grepl("fentanyl", text)) %>%
  group_by(date) %>%
  summarise(methadone_mentions = sum(methadone),
            suboxone_mentions = sum(suboxone),
            fentanyl_mentions = sum(fentanyl))

# Reshape data for plotting
mentions_long <- S2_mentions_by_date %>%
  pivot_longer(cols = c(methadone_mentions, suboxone_mentions, fentanyl_mentions),
               names_to = "substance", values_to = "mentions")

# Plot the time-series comparison
ggplot(mentions_long, aes(x = date, y = mentions, color = substance)) +
  geom_line(size = 1) +
  labs(title = "A1_set2 Mentions of Methadone, Suboxone, and Fentanyl Over Time",
       x = "Date", y = "Number of Mentions",
       color = "Substance") +
  theme_minimal()
```

8. Find the top 10 most frequent bigrams in each of the three sets. Plot a bar chart for these.

```{r}
ignore_words <- c("https", "t.co", "methadone", "suboxone", "fentanyl", 
                  "methadone_cat", "amp", "danlairdmd", "tootwisted40", "2", 
                  "is", "i", "a", "the", "to", "of", "in", "and", "if", "it")

# Function to extract and count bigrams, while removing ignore words
top_bigrams <- function(data, substance) {
  data %>%
    filter(grepl(substance, text)) %>%                               # Filter for posts mentioning the substance
    unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%         # Tokenize the text into bigrams
    separate(bigram, into = c("word1", "word2"), sep = " ") %>%      # Separate bigrams into two columns
    filter(!word1 %in% ignore_words, !word2 %in% ignore_words) %>%   # Remove bigrams with ignore words
    unite(bigram, word1, word2, sep = " ") %>%                       # Recombine the bigrams
    count(bigram, sort = TRUE) %>%                                   # Count occurrences of each bigram
    top_n(10, n)                                                     # Get the top 10 most frequent bigrams
}

# Extract top bigrams for each substance
bigrams_methadone <- top_bigrams(A1_set1, "methadone")
bigrams_suboxone <- top_bigrams(A1_set1, "suboxone")
bigrams_fentanyl <- top_bigrams(A1_set1, "fentanyl")

# Plot function for bigrams
plot_bigrams <- function(bigrams_data, substance_name) {
  ggplot(bigrams_data, aes(x = reorder(bigram, n), y = n)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() +
    labs(title = paste("Top 10 A1_set1 Bigrams Associated with", substance_name),
         x = "Bigrams", y = "Frequency") +
    theme_minimal()
}

# Plot top A1_set1 bigrams for methadone
plot_bigrams(bigrams_methadone, "Methadone")

# Plot top A1_set1 bigrams for suboxone
plot_bigrams(bigrams_suboxone, "Suboxone")

# Plot top A1_set1 bigrams for fentanyl
plot_bigrams(bigrams_fentanyl, "Fentanyl")
```

```{r}
# Function to extract and count bigrams, while removing ignore words
top_bigrams <- function(data, substance) {
  data %>%
    filter(grepl(substance, text)) %>%
    unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
    separate(bigram, into = c("word1", "word2"), sep = " ") %>%
    filter(!word1 %in% ignore_words, !word2 %in% ignore_words) %>%
    unite(bigram, word1, word2, sep = " ") %>%
    count(bigram, sort = TRUE) %>%
    top_n(10, n)
}

# Extract top bigrams for each substance
bigrams_methadone_s2 <- top_bigrams(A1_set2, "methadone")
bigrams_suboxone_s2 <- top_bigrams(A1_set2, "suboxone")
bigrams_fentanyl_s2 <- top_bigrams(A1_set2, "fentanyl")

# Plot function for bigrams
plot_bigrams <- function(bigrams_data, substance_name) {
  ggplot(bigrams_data, aes(x = reorder(bigram, n), y = n)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() +
    labs(title = paste("Top 10 A1_set2 Bigrams Associated with", substance_name),
         x = "Bigrams", y = "Frequency") +
    theme_minimal()
}

# Plot top A1_set1 bigrams for methadone
plot_bigrams(bigrams_methadone_s2, "Methadone")

# Plot top A1_set1 bigrams for suboxone
plot_bigrams(bigrams_suboxone_s2, "Suboxone")

# Plot top A1_set1 bigrams for fentanyl
plot_bigrams(bigrams_fentanyl_s2, "Fentanyl")
```